---
title: "Discussion session"
author: "BJ"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
library(here)
library(tidyverse)
library(sf)
library(tmap)
```

```{r}
sb_protected_areas <- read_sf(here::here("data", "cpad_super_units_sb.shp")) %>% 
  st_transform("ESRI:102009")

sb_city_boundaries <- read_sf(here::here("data", "sb_city_boundaries_2003.shp")) %>%
  st_transform("ESRI:102009")

sb_county_boundary <- read_sf(here::here("data", "sb_county_boundary_2020.shp")) %>%
  st_transform("ESRI:102009")

aves <- read_sf(here::here("data", "aves_observations_2020_2024.shp")) %>%
  st_transform("ESRI:102009")
```

Now, to meet this week’s learning objectives, your task:

- Find how many bird observations are within protected areas in Santa Barbara County
```{r}
aves_PA_subset <- sb_protected_areas[aves, ]
nrow(aves_PA_subset)
```

```{r}
tm_shape(sb_protected_areas) +
  tm_polygons() +
  tm_shape(aves_PA_subset) +
  tm_dots()
```


```{r}
aves_PA_join <- st_join(aves, sb_protected_areas)
nrow(aves_PA_join)
```


```{r}
tm_shape(aves_PA_subset) +
  tm_polygons() +
  tm_shape(aves_PA_join) +
  tm_dots()
```


```{r}
st_crs(sb_protected_areas)$units
```

```{r}
PAs_5km <- st_buffer(sb_protected_areas, dist = 500)
```

```{r}
aves_PAs_5km <- PAs_5km[aves, ]
nrow(aves_PAs_5km)
```


```{r}
goleta <- sb_city_boundaries |>
  filter(NAME == "Goleta")

goleta_5km <- st_buffer(sb_protected_areas, dist = 500)
```

```{r}

```


- Show the different outputs from a spatial subset and a spatial join

- Bonus Challenge: Try it out with a 5 km buffer around the protected areas too!

Find the protected areas within 15 km of a city in Santa Barbara County

Hint: Use dplyr::filter() to select a city from sb_city_boundaries

Explore the different outputs with st_intersects(), st_intersection(), and st_within()

Practice a distance-based join with st_is_within_distance()

Find the distance between your city of choice and a protected area of your choice

Note: st_distance() finds the distance between the geometries’ edges